Radial and rotational velocities of young brown dwarfs and very low-mass stars in the Upper Scorpius OB association and the $\rho$~Ophiuchi cloud core

### 2006-05-08 BD_radvel/kurosawa.tex
minor change
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@712 25464f21-6099-402c-b30a-640f1d40d38b

We present the results of a radial velocity (RV) survey of 14 brown
dwarfs (BDs) and very low-mass (VLM) stars in the Upper Scorpius OB
association (UScoOB) and 3 BD candidates in the $\rho$~Ophiuchi dark
cloud core. We obtained high-resolution echelle spectra at VLT using
Ultraviolet and Visual Echelle Spectrograph (UVES) at two different
epochs for each object, and measured the shifts in their RVs to
identify candidates for binary/multiple systems in the sample. The
average time separation of the RV measurements is 21.6~d, and our
survey is sensitive to the binaries with separation $<0.12$~au.  We
found 2 out of 17 objects (or $12^{+13}_{-10}$~per~cent by fraction)
show a significant RV change in 4--33~d time scale, and are 
considered as binary/multiple `candidates.'  We found no double-lined
spectroscopic binaries in our sample, based on the shape of
cross-corelation curves. The RV dispersion of the objects in UScoOB
is found to be very similar to that of the BD and VLM stars in
Chamaeleon~I (Cha~) previously studies by others. We also found the
distribution of the mean rotational velocities ($v\,\sin i$) of the
UScoOB objects is similar to that of the Cha~I, but the dispersion of
$v\,\sin i$ is much larger than that of the Cha~I objects.

@@

The extensive imaging surveys provide excellent observational constraints
on wider BD+BD binaries, but it is now necessary to search for shorter
period BD+BD binaries systematically. Binaries with the separation
of less than 1~au are not resolved by current imaging techniques,
but will be detectable as spectroscopic binaries, providing the mass-ratio
is not too extreme, and velocity separation is large enough. The first
BD+BD spectroscopic binary, PPl~15  \citep{basri:1999} showed a
double-peaked cross-correlation function with a maximum velocity separation
of $>70\,\kmps$. The binary was found to have an eccentric orbit
($e=0.4$) with a period of $\sim5.8\,\mathrm{d}$. \citet{basri:1999}
suggested that the formation process of substellar objects is biased
towards smaller separation binaries based on the short period of PPl~15
and the lack of Pleiades BD binaries with separations $>40\,\mathrm{au}$.
Note that the median separation of binaries with solar-type primaries
is 30~au \citet{duquennoy:1991}. Pioneering work on the RVs
of BDs and VLMs are presented by \citet{guenther:2003}, \cite{kenyon:2005}
and \citet{joergens:2006} who found a several binary candidates;
however, the orbital parameters and masses of binaries remains
unknown because the follow-up spectroscopic monitoring is lacking
or still being undertaken. In addition to the follow-up observations,
the number of BDs and VLM star binary candidates needs to be increased
in order to have better statistics on short-period binary parameters.

@@

Before applying the cross-correlation technique to out main targets,
we have applied the technique to the radial velocity standard HD~140538
for which an high-accuracy RV measurement via the fixed-configuration,
cross-dispersed échelle spectrograph Elodie \citep{baranne:1996}
is available. This was done so to ensure not only the validity of
the cross-correlation technique, but also the validity of the wavelength
calibration. In this test, we found the heliocentric $\mathrm{RV_{HD\,140538}=}18.8\pm0.6\,\kmps$
which is in good agreement with the Elodie radial velocity measurement
of $19.00\pm0.05\,\mathrm{km\, s^{-1}}$ \citep{udry:1999}.

The result of the heliocentric RV measurements (from two epochs for
each object) is summarised in Table~\ref{tab:summary_results} along
with the uncertainties. The table also lists the uncertainties in
the `relative' radial velocities ($\sigma_{\mathrm{RRV}}$) which
is the uncertainty of RV with respect to the template star. The
two consecutive measurements of RVs (from same nights) are used to
find the (absolute/heliocentric) RV of the night, and their uncertainties are obtained from
the standard deviations of the mean ($\sigma_{\mathrm{RRV}}$) and
the uncertainty of the template RV value (c.f.~\citealt{joergens:2006}).
Note that for finding a shift in the radial velocities from two
different epochs, only $\sigma_{\mathrm{RRV}}$ is important since one
only requires relative (with respect to a template) RV values.
The 'average' radial velocities $\left (\overline{\mathrm{RV}} \right )$ of the
two epochs are also given in the same table. The $\overline{\mathrm{RV}}$
values of the objects in UScoOB are consistent with the earlier measurements
of \citet{muzerolle:2003} in Table~\ref{tab:literatures}.

For each object and for each RV measurement, the deviations ($\Delta\mathrm{RV}$)
from the average RV are computed and summarized in Fig.~\ref{fig:radvel01}
along with their uncertainties ($\sigma_{\mathrm{RRV}}$) in order
to aid the identification of multiplicity candidates. Note that in
computing $\Delta\mathrm{RV}$ we do not require the knowledge of
absolute or the heliocentric radial velocities, but only the relative
velocities (with respect to a template).

@@

We have computed $p$ for all the objects, and have plotted the results 
as a histogram of the $-\log{p}$, shown in Fig.~\ref{fig:chisq-test}
(excluding the non-member USco~121; see explanation later). 
The figure clearly shows two distinctive populations: one on the left
(with small $-\log{p}$ values) occupied by the RV constant objects,
and one on in the right-most bin (with the largest $-\log{p}$)
occupied by the RV variable objects which have $p<10^{-3}$.  The
expected distribution of $p$ computed consistently with our
uncertainty measurements is also shown in the same figure. The RV
constant population on the left side reasonably matches the expected
curve. This assures that our uncertainties in RV values are
reasonable, and an additional systematic error may not be necessary in
this analysis (but see Section~\ref{sec:Binary-fraction}). According
to the figure, there are four objects which show significant RV
variations out of 17 samples. They are GY~141, GY~310, USco~55, and
USco~104, and considered as our preliminary binary/multiple
candidates.  We will discuss the effect of the systematics error in detail, and
list our final binary/multiple candidate later in
Section~\ref{sec:Binary-fraction}.

@@

The range of $v\,\sin i$ found among our objects is $3.6$--$55.6\,\kmps$,
and a similar range is also found by \citet{mohanty:2005}. 
Fig.~\ref{fig:vrot_histrogram} shows the histogram
of $v\,\sin i$ distribution for the UScoOB objects (14 objects excluding
USco 121, non member). The log-normal fit of this distribution gives
the peak position at $16.9\,\kmps$ with a standard deviation $\sigma=27.8\,\kmps$.
Using the $v\,\sin i$ data in \citet{joergens:2001}, the same histogram
bin size used for UScoOB obects and the log-normal fit, we find the
$v\,\sin i$ distribution of the BD and VLM stars (8 objects) in Cha~I
peaks at $15.4\,\kmps$, and has the standard deviation of $8.0\,\kmps$.
The peak of the distribution is similar to that of Upper Sco objects,
but the standard deviation of the distribution is significantly smaller
than that of the Upper Sco objects. The difference maybe due to the
very small sample. A similar fit was applied to the $v\,\sin i$
distribution of 14 T~Tauri stars in Cha~I using the data of \citet{joergens:2001},
and we found a peak at $17.0\,\kmps$ with a  standard deviation
$25.9\,\kmps$ which are very similar to those of the Upper Sco brown
dwarf candidates and VLM stars.

>>>>

We present the results of a radial velocity (RV) survey of 14 brown
dwarfs (BDs) and very low-mass (VLM) stars in the Upper Scorpius OB
association (UScoOB) and 3 BD candidates in the $\rho$~Ophiuchi dark
cloud core. We obtained high-resolution echelle spectra at VLT using
Ultraviolet and Visual Echelle Spectrograph (UVES) at two different
epochs for each object, and measured the shifts in their RVs to
identify candidates for binary/multiple systems in the sample. The
average time separation of the RV measurements is 21.6~d, and our
survey is sensitive to the binaries with separation $<0.12$~au.  We
found 2 out of 17 objects (or $24^{+16}_{-14}$~per~cent by fraction)
show a significant RV change in 4--33~d time scale, and are 
considered as binary/multiple `candidates.'  We found no double-lined
spectroscopic binaries in our sample, based on the shape of
cross-corelation curves. The RV dispersion of the objects in UScoOB
is found to be very similar to that of the BD and VLM stars in
Chamaeleon~I (Cha~) previously studies by others. We also found the
distribution of the mean rotational velocities ($v\,\sin i$) of the
UScoOB objects is similar to that of the Cha~I, but the dispersion of
$v\,\sin i$ is significantly larger than that of the Cha~I objects.

@@

The extensive imaging surveys provide excellent observational constraints
on wider BD+BD binaries, but it is now necessary to search for shorter
period BD+BD binaries systematically. Binaries with the separation
of less than 1~au are not resolved by current imaging techniques,
but will be detectable as spectroscopic binaries, providing the mass-ratio
is not too extreme, and velocity separation is large enough. The first
BD+BD spectroscopic binary, PPl~15  \citep{basri:1999}, showed a
double-peaked cross-correlation function with a maximum velocity separation
of $>70\,\kmps$. The binary was found to have an eccentric orbit
($e=0.4$) with a period of $\sim5.8\,\mathrm{d}$. \citet{basri:1999}
suggested that the formation process of substellar objects is biased
towards smaller separation binaries based on the short period of PPl~15
and the lack of Pleiades BD binaries with separations $>40\,\mathrm{au}$.
Note that the median separation of binaries with solar-type primaries
is 30~au \citet{duquennoy:1991}. Pioneering work on the RVs
of BDs and VLMs are presented by \citet{guenther:2003}, \cite{kenyon:2005}
and \citet{joergens:2006} who found a several binary candidates;
however, the orbital parameters and masses of binaries remains
unknown because the follow-up spectroscopic monitoring is lacking
or still being undertaken. In addition to the follow-up observations,
the number of BDs and VLM star binary candidates needs to be increased
in order to have better statistics on short-period binary parameters.

@@

Before applying the cross-correlation technique to our main targets,
we have applied the technique to the radial velocity standard HD~140538
for which an high-accuracy RV measurement via the fixed-configuration,
cross-dispersed échelle spectrograph Elodie \citep{baranne:1996}
is available. This was done so to ensure not only the validity of
the cross-correlation technique, but also the validity of the wavelength
calibration. In this test, we found the heliocentric $\mathrm{RV_{HD\,140538}=}18.8\pm0.6\,\kmps$
which is in good agreement with the Elodie radial velocity measurement
of $19.00\pm0.05\,\mathrm{km\, s^{-1}}$ \citep{udry:1999}.

The result of the heliocentric RV measurements (from two epochs for
each object) is summarised in Table~\ref{tab:summary_results} along
with the uncertainties. The table also lists the uncertainties in
the `relative' radial velocities ($\sigma_{\mathrm{RRV}}$) which
is the uncertainty of RV with respect to the template star. The
two consecutive measurements of RVs (from same nights) are used to
find the (absolute/heliocentric) RV of the night, and their uncertainties are obtained from
the standard deviations of the mean ($\sigma_{\mathrm{RRV}}$) and
the uncertainty of the template RV value (c.f.~\citealt{joergens:2006}).
Note that for finding a shift in the radial velocities from two
different epochs, only $\sigma_{\mathrm{RRV}}$ is important since one
only requires relative (with respect to a template) RV values.
The `average' radial velocities $\left (\overline{\mathrm{RV}} \right )$ of the
two epochs are also given in the same table. The $\overline{\mathrm{RV}}$
values of the objects in UScoOB are consistent with the earlier measurements
of \citet{muzerolle:2003} in Table~\ref{tab:literatures}.

For each object and for each RV measurement, the deviations ($\Delta\mathrm{RV}$)
from the average RV are computed and summarized in Fig.~\ref{fig:radvel01}
along with their uncertainties ($\sigma_{\mathrm{RRV}}$) in order
to aid the identification of multiplicity candidates. Note that in
computing $\Delta\mathrm{RV}$ we do not require the knowledge of
absolute radial velocities, but only the relative
velocities (with respect to a template).

@@

We have computed $p$ for all the objects, and have plotted the results 
as a histogram of the $-\log{p}$, shown in Fig.~\ref{fig:chisq-test}
(excluding the non-member USco~121; see explanation later). 
The figure clearly shows two distinctive populations: one on the left
(with small $-\log{p}$ values) occupied by the RV constant objects,
and one on in the right-most bin (with the largest $-\log{p}$)
occupied by the RV variable objects which have $p<10^{-3}$.  The
expected distribution of $p$ computed consistently with our
uncertainty measurements is also shown in the same figure. The RV
constant population on the left side reasonably matches the expected
curve. This reassures that our uncertainties in RV values are
reasonable, and an additional systematic error may not be necessary in
this analysis (but see Section~\ref{sec:Binary-fraction}). According
to the figure, there are four objects which show significant RV
variations out of 17 samples. They are GY~141, GY~310, USco~55, and
USco~104, and considered as our preliminary binary/multiple
candidates.  We will discuss the effect of the systematics error in detail, and
list our final binary/multiple candidates in Section~\ref{sec:Binary-fraction}.

@@

The range of $v\,\sin i$ found among our objects is $3.6$--$55.6\,\kmps$,
and a similar range is also found by \citet{mohanty:2005}. 
Fig.~\ref{fig:vrot_histrogram} shows the histogram
of $v\,\sin i$ distribution for the UScoOB objects (14 objects excluding
USco 121, non member). The log-normal fit of this distribution gives
the peak position at $16.9\,\kmps$ with a standard deviation $\sigma=27.8\,\kmps$.
Using the $v\,\sin i$ data in \citet{joergens:2001}, the same histogram
bin size used for UScoOB obects and the log-normal fit, we find the
$v\,\sin i$ distribution of the BD and VLM stars (8 objects) in Cha~I
peaks at $15.4\,\kmps$, and has the standard deviation of $8.0\,\kmps$.
The peak of the distribution is similar to that of Upper Sco objects,
but the standard deviation of the distribution is significantly smaller
than that of our Upper Sco objects. The difference maybe due to the
very small sample. A similar fit was applied to the $v\,\sin i$
distribution of 14 T~Tauri stars in Cha~I using the data of \citet{joergens:2001},
and we found a peak at $17.0\,\kmps$ with a  standard deviation
$25.9\,\kmps$ which are very similar to those of the Upper Sco brown
dwarf candidates and VLM stars.
